# Descriptive statistics {#descriptives}
This section will give you a brief overview of how to generate descriptive statistics in R. 

>Note: For this section and all subsequent section, use `?` in front of a function name to retrieve help files on this function.

## Inspect the dataframe
To see collumn names, use the `colnames()` function:
```{r eval=FALSE}
# See variable names
# GLOBE
colnames(globe_data)
# RECORD
colnames(record_data)
```

The `head()` function gives you the variable names, types, and the 6 first lines of values in a dataset.
```{r eval=FALSE}
# Inspect GLOBE and RECORD datasets
# GLOBE
head(globe_data)
# RECORD
head(record_data)
```

## Measures of central tendancy and variability
Basic R functions for computing descriptive statistics include:
 
 * `mean()` - returns the mean of a numeric variable
 * `sd()` - returns the standard deviation of a numeric variable
 * `var()` - returns de median of a numeric variables
 * `min()` - returns de minimum of a numeric variables
 * `maximum()` - returns de maximum of a numeric variables
 * `median()`- returns de median of a numeric variables
 * `range()` - returns the range of values of a numeric variables (minimum and maximum)
 * `IQR()` - returns the interquartile range of a numeric variables
 
```{r eval=FALSE}
# Mean age for GLOBE baseline collection wave 
mean(globe_data$sdc_age_0)
# Mean age for RECORD baseline collection wave 
mean(record_data$sdc_age_0)
```
 
The `summary()` command returns several desciptive statistics summaries for numeric variables. you can either run this command for a single variable or for an entire dataframe. For a single variable, this function return the mean, median, 25th and 75th quartiles, minimum and maximum values. When applied to a dataframe, this command generates these stastitics for all collumns in this dataframe. 

```{r eval=FALSE}
# Summary of the GLOBE baseline collection wave age variable
summary(globe_data$sdc_age_0)
# Summary of the entire GLOBE dataframe
summary(globe_data)
```

The `summ()` function from the `epiDisplay` package is also quite usefull. In addition to the number of observations, mean, median, standard deviation, minimum, maximum, and number of missing values, it also generates a plot showing the distribution of the variable. 


## Categorical variables
For factor variables (i.e. categorical variables), the `table()` function returns the number of observation per category. 
```{r eval=FALSE}
# Number of Males/Females in GLOBE
table(globe_data$sdc_gender_0)
# Number of Males/Females in RECORD
table(record_data$sdc_gender_0)
```

Wrapping this function in prop.table() give the proportion of Males and Females in each study.
```{r eval=FALSE}
# Proportion of Males/Females in GLOBE
prop.table(table(globe_data$sdc_gender_0))
# Proporion of Males/Females in RECORD
prop.table(table(record_data$sdc_gender_0))
```

Again, the epiDisplay package provides a useful function to summarise categorical data. Running the `tab1()` function gives you the frequency and proportion for each category and generates a histogram for a variable. Missing data are also shown in tables and histograms.
```{r eval=FALSE}
# Proportion of Males/Females in GLOBE
tab1(globe_data$sdc_gender_0)
# Proporion of Males/Females in RECORD
tab1(record_data$sdc_gender_0)
```

## Graphs and plots
R also allows you to generate graphics and plots to visualise the harmonized data. For example, you can create a box plot using the `boxplot()` function or a histogram using the `hist()` function.
```{r eval=FALSE}
# Box plot and histogram of age at baseline for GLOBE Study
boxplot(globe_data$sdc_age_0)
hist(record_data$sdc_age_0)

# Box plot and histogram of age at baseline for RECORD
boxplot(record_data$sdc_age_0)
hist(record_data$sdc_age_0)
```
>Note: access the help files using `?boxplot` and `?hist` to learn how to format your box plot or histogram, including how to create a title for your graph, label axes, add colour or subset observations or apply breaks to numerical data.


## Pooled statistics
For pooled analyses, we will need  a study indicator variable that will adjust for study effect in regression models (Section \@ref(longitudinal)) . The following script creates a new collumn in each dataset called _study_ and gives it a value of 'RECORD' for all observations in record_data and GLOBE for all observations in globe_data.
```{r eval=FALSE}
# Create a new column called study and assign values of 0 (for RECORD) and 1 (for GLOBE)
record_data['study']= "RECORD"
globe_data['study']= "GLOBE"
```

Next, we use the `rbind()` command to combine globe_data and record_data into a single pooled dataset.
```{r eval=FALSE}
# Combine dataframes into a pooled dataset using the rbind command
pooled_data <- rbind(globe_data, record_data)
```

You can play around with the functions outlined above to get descriptive statistics for the pooled dataset.
