# Cross-sectional analyses {#cross-sectional}
In this section, simple examples of cross-sectional linear and logistic regressions for single studies and for pooled data are presented.

## Linear regression
The `glm()` function allows fitting a generalized linear regression model. The example below is a corss-sectional (baseline assessment data) linear regression with body mass index as the outcome and age, sex, education level, household income and smoking status as predictor variables. 
```{r eval=FALSE}
# GLM for BMI using GLOBE study
glm_bmi <- glm(phy_bmi_all_0 ~ sdc_age_0 + sdc_gender_0 + sdc_highest_edu_0 + sdc_income_0 + lsb_smk_status_0, family = gaussian, data = globe_data)
```

Using the `<-` assignment operator allows us to store the results of this analysis in an R object called glm_bmi. Applying the `summary()` function on this object allows us to see the results. 

```{r eval=FALSE}
# See linear regression results
summary(glm_bmi)
```

The `glm()` function can also be applied to a categorical outcome by specifi

## Logistic regression
For 
