# Longitudinal analyses {#longitudinal}
This section provides simple examples of how to go about conducting longitudinal analyses of harmonized MINDMAP data.

## Wide to long format
Before we conduct longitudinal analyses, we need to transform each dataframe (GLOBE and RECORD) from wide to long format. The following scripts does this transformation for each study and saves the long-format dataframes as *globe_data_long* and *record_data_long*.
```{r eval=FALSE}
# GLOBE: wide to long format
globe_data_long <- globe_data %>%
  gather(var, value, -entity_id, -study) %>%
  separate(var, into = c("variable", "time"), sep = ("_(?=[^_]+$)")) %>%
  spread(variable, value)

# RECORD: wide to long format
record_data_long <- record_data %>%
  gather(var, value, -entity_id, -study) %>%
  separate(var, into = c("variable", "time"), sep = ("_(?=[^_]+$)")) %>%
  spread(variable, value)
```

Next, we combine GLOBE and RECORD long format dataframes into a pooled dataframe:
```{r eval=FALSE}
# Pooled data: long format
pooled_data_long <- rbind(globe_data_long, record_data_long)
```
>Note: In the wide to long format transformation, the *_0* and *_1* suffixes, indicating baseline and follow up data collections are removed from the datasets.

## Linear mixed models
### Continuous outcome (normally distributed)
The `lmer()` function from the `lme4` allows us to fit linear mixed-effect models. In the example below, data for two time points (baseline and follow up 1) of the GLOBE study are fitted with body mass index as the outcome and age, sex, household income and smoking status as predictor variables. 
```{r eval=FALSE}
lmm_bmi_globe <- lmer(phy_bmi_all ~ sdc_age + sdc_gender + sdc_income + lsb_smk_status + (1 | entity_id), data = globe_data_long)

summary(lmm_bmi_globe)
```

The results in *lmm_bmi_globe* do not include p-values. To get p-values, the `Anova()` function can be used:
```{r eval=FALSE}
Anova(lmm_bmi_globe)
```

Below, the same model is fitted to the pooled dataset adding the study indicator variable to the formula.
```{r eval=FALSE}
lmm_bmi_pooled <- lmer(phy_bmi_all ~ sdc_age + sdc_gender + sdc_income + lsb_smk_status + study + (1 | entity_id), data = pooled_data_long)

summary(lmm_bmi_pooled)

Anova(lmm_bmi_pooled)
```

### Binary outcome

For binary outcomes, the `glmer()` function is used to fit a generalized linear mixed-effects model. The example below 
```{r eval=FALSE}
lmm_bmi_pooled_binary <- glmer(poh_sah_poor ~ sdc_age + sdc_gender + sdc_income + lsb_smk_status + study + (1 | entity_id), data = pooled_data_long_2, family = binomial)
```
You will get a _Convergence warning_ message. The following webpage provides troubleshooting for such problems: 
<https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html>

In the example above, the *sdc_age* variable needs to be scalled. To do this, we used the `scale()` function and save the scalling resuls as a new variable called *sdc_age_sc*: 
```{r eval=FALSE}
pooled_data_long$sdc_age_sc <- scale(pooled_data_long$sdc_age) 
```

This variable then replaces the original *sdc_age* variable:
```{r eval=FALSE}
lmm_bmi_pooled_binary <- glmer(poh_sah_poor ~ sdc_age_sc + sdc_gender + sdc_income + lsb_smk_status + study + (1 | entity_id), data = pooled_data_long, family = binomial)

summary(lmm_bmi_pooled_binary)
```

