# Longitudinal analyses {#longitudinal}
This section provides simple examples of how to go about conducting longitudinal analyses of harmonized MINDMAP data.

## Wide to long format
Before we conduct longitudinal analyses, we need to transform each dataframe (GLOBE and RECORD) from wide to long format. The following scripts does this transformation for each study and saves the new dataframes as *globe_data_long* and *record_data_long*.
```{r eval=FALSE}
# GLOBE: wide to long format
globe_data_long <- globe_data %>%
  gather(var, value, -entity_id, -study) %>%
  separate(var, into = c("variable", "time"), sep = ("_(?=[^_]+$)")) %>%
  spread(variable, value)

# RECORD: wide to long format
record_data_long <- record_data %>%
  gather(var, value, -entity_id, -study) %>%
  separate(var, into = c("variable", "time"), sep = ("_(?=[^_]+$)")) %>%
  spread(variable, value)
```

Next, we combine GLOBE and RECORD long format dataframes into a pooled dataframe:
```{r eval=FALSE}
# Pooled data: long format
pooled_data_long <- rbind(globe_data_long, record_data_long)
```
>Note: in the *long format* dataframes the *_0* and *_1* suffixes, indicating baseline and follow up data collections are removed from the datasets.

## Linear mixed models
### Continuous outcome (normally distributed)
The `lmer()` function from the `lme4` allows us to fit linear mixed-effect models. In the example below, data for two time points (baseline and follow up 1) of the GLOBE study are fitted with body mass index as the outcome and age, sex, household income and smoking status as predictor variables. 
```{r eval=FALSE}
lmm_bmi_globe <- lmer(phy_bmi_all ~ sdc_age + sdc_gender + sdc_income + lsb_smk_status + (1 | entity_id), data = globe_data_long)

summary(lmm_bmi_globe)
```

Below, the same model is fitted to the pooled dataset adding the study indicator variable to the formula.
```{r eval=FALSE}
lmm_bmi_pooled <- lmer(phy_bmi_all ~ sdc_age + sdc_gender + sdc_income + lsb_smk_status + study + (1 | entity_id), data = pooled_data_long)

summary(lmm_bmi_pooled)
```

### Binary outcome

For binary outcomes, the `glmer()` function is used to fit a generalized linear mixed-effects model.
```{r eval=FALSE}
lmm_bmi_pooled <- lmer(phy_bmi_all ~ sdc_age + sdc_gender + sdc_income + lsb_smk_status + study + (1 | entity_id), data = pooled_data_long)

summary(lmm_bmi_pooled)
```

