--- 
title: "MINDMAP data platform tutorial"
author: "Dany Doiron"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a short tutorial on how to use the MINDMAP data platform in RStudio."
---

# Tutorial overview

Using harmonized data from two MINDMAP studies, participants will learn:
* How to access the harmonized data on the Opal server via the Web RStudio interface
* How to manipulate harmonized data in the R environment
* How to generate study specific and pooled descriptive statistics in the RStudio environment
* How to run study specific and pooled regression analyses in the R environment

The `install.packages()` command allows you to install R packages for your analyses. Run the code below to install packages needed for this tutorial

```{r eval=FALSE}
# Install and load required libraries for this tutorial
install.packages("epiDisplay")
install.packages("tidyverse")
library(epiDisplay)
library(tidyverse)
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
